Printing in a Periphery: a Quantitative Study of Finnish Knowledge Production, 1640-1828

### 2016-07-12 inst/examples/20160715-slides-Krakow.Rmd
sync
s

### Data processing is a major effort and should not be underestimated

We have built custom tools to automate all analyses from raw data to final summary documents (these slides, for instance).

\begin{columns}
\begin{column}{0.5\textwidth}
     \includegraphics[width=1\textwidth]{parsing.png}
\end{column}
\begin{column}{0.5\textwidth}  %%<--- here
    \begin{center}
     \includegraphics[width=1\textwidth]{iceberg.png}
     \end{center}
\end{column}
\end{columns}

![](ecosystemflowchart.png)

@@

### Practical exercise: Quantitative Study of Finnish Knowledge Production, 1640-1828

@@

1. General trends & political events

@@

```{r publishingactivitycomparisons, echo=FALSE, message=FALSE, cache=TRUE, fig.width=8, fig.height=5, fig.show="hold", out.width="150px"}
v <- seq(1500, 1800, 100)
selected.places = c("Turku", "Stockholm")
df2 <- df0 %>% group_by(publication_decade, publication_place, catalog) %>%
               summarise(n = n()) %>%
	       filter(publication_place %in% selected.places) 
df2$catalog = factor(df2$catalog, levels = rev(c("Fennica", "Kungliga")))
df2$publication_place = droplevels(factor(df2$publication_place))
df3 = as.data.frame(spread(df2, publication_decade, n, fill = 0))
df2 = melt(df3)
colnames(df2) = c("publication_place", "catalog", "publication_decade", "n")
df2$publication_decade = as.numeric(as.character(df2$publication_decade))
df2$n = as.numeric(as.character(df2$n))
for (catal in unique(df2$catalog)) {
  p <- ggplot(subset(df2, catalog == catal), aes(x = publication_decade, y = n)) +
     geom_line(aes(linetype = publication_place)) +
     geom_point(aes(shape = publication_place), size = 3) +
     ylab("Title count (n)") + xlab("Publication year") +  
     ggtitle(catal) +
     scale_x_continuous(breaks = v, labels = v) + 
     theme(axis.text.x = element_text(angle = 0))
  print(p)

@@

2. Publishers in Sweden and Turku

@@

3. Languages & religious texts in Finland

@@

4. University curriculum & Enlightenment in Finland

@@

```{r LIBER-13, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=10, fig.height=8, fig.show="hold", out.width="150px"}
for (catal in unique(df0$catalog)) {
  df2 <- subset(df0, catalog == catal) %>% group_by(publication_decade, gatherings) %>% summarize(paper.consumption.km2 = sum(paper.consumption.km2, na.rm = TRUE), n = n()) 
  df2 <- filter(df2, gatherings %in% setdiff(names(which(table(df2$gatherings) >= 15)), "NA"))
  df2$highlight <- rep("gray", nrow(df2))
  df2$highlight[df2$gatherings == "8vo"] <- "darkgreen"

@@

```{r authors22, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6, out.width="160px", fig.show="hold"}
theme_set(theme_bw(20))
for (catalogue in c("Fennica", "Kungliga")) {
  df <- filter(df0, catalog == catalogue) %>%
        filter(publication_year >= 1640 & publication_year <= 1828)
  top.authors <- names(bibliographica::top(df, field = "author", n = 10))
  dfs <- df %>% filter(author %in% top.authors) %>%
         group_by(author, publication_decade) %>%
         tally() %>%
         arrange(publication_decade)
  p <- ggplot(dfs, aes(x = publication_decade, y = n, fill = author)) +
       geom_bar(stat = "identity", position = "stack", color = "black") +
       xlab("Publication Decade") +
       ylab("Title Count") +
       scale_fill_grey() +
       guides(fill = guide_legend("Author")) +
       ggtitle(paste("Title count top authors\n", catalogue)) +
       xlim(c(1640, 1828))
  print(p)
}
```

>>>>

@@

### Outline of Quantitative Study of Finnish Knowledge Production, 1640-1828

@@

**1. General trends & political events**

@@

```{r publishingactivitycomparisons, echo=FALSE, message=FALSE, cache=TRUE, fig.width=10, fig.height=9, fig.show="hold", out.width="160px"}
v <- seq(1500, 1800, 100)
selected.places = c("Turku", "Stockholm")
df2 <- df0 %>% group_by(publication_decade, publication_place, catalog) %>%
               summarise(n = n()) %>%
	       filter(publication_place %in% selected.places) 
df2$catalog = factor(df2$catalog, levels = rev(c("Fennica", "Kungliga")))
df2$publication_place = droplevels(factor(df2$publication_place))
df3 = as.data.frame(spread(df2, publication_decade, n, fill = 0))
df2 = melt(df3)
colnames(df2) = c("publication_place", "catalog", "publication_decade", "n")
df2$publication_decade = as.numeric(as.character(df2$publication_decade))
df2$n = as.numeric(as.character(df2$n))
for (catal in unique(df2$catalog)) {
  p <- ggplot(subset(df2, catalog == catal), aes(x = publication_decade, y = n)) +
     geom_line(aes(linetype = publication_place)) +
     geom_point(aes(shape = publication_place), size = 3) +
     ylab("Title count (n)") + xlab("Publication year") +  
     ggtitle(catal) +
     scale_x_continuous(breaks = v, labels = v) + 
     theme(axis.text.x = element_text(angle = 0))
  print(p)

@@

**2. Publishers in Sweden and Turku**

@@

**3. Languages & religious texts in Finland**

@@

**4. University curriculum & Enlightenment in Finland**

@@

```{r LIBER-13, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=11, fig.height=9, fig.show="hold", out.width="160px"}
for (catal in unique(df0$catalog)) {
  df2 <- subset(df0, catalog == catal) %>% group_by(publication_decade, gatherings) %>% summarize(paper.consumption.km2 = sum(paper.consumption.km2, na.rm = TRUE), n = n()) 
  df2 <- filter(df2, gatherings %in% setdiff(names(which(table(df2$gatherings) >= 15)), "NA"))
  df2$highlight <- rep("gray", nrow(df2))
  df2$highlight[df2$gatherings == "8vo"] <- "darkgreen"

@@

```{r authors22, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=10, out.width="160px", fig.show="hold"}
theme_set(theme_bw(20))
for (catalogue in c("Fennica", "Kungliga")) {
  df <- filter(df0, catalog == catalogue) %>%
        filter(publication_year >= 1640 & publication_year <= 1828)
  top.authors <- names(bibliographica::top(df, field = "author", n = 10))
  dfs <- df %>% filter(author %in% top.authors) %>%
         group_by(author, publication_decade) %>%
         tally() %>%
         arrange(publication_decade)
  p <- ggplot(dfs, aes(x = publication_decade, y = n, fill = author)) +
       geom_bar(stat = "identity", position = "stack", color = "black") +
       xlab("Publication Decade") +
       ylab("Title Count") +
       scale_fill_grey() +
       guides(fill = guide_legend("Author")) +
       ggtitle(paste("Title count top authors\n", catalogue)) +
       xlim(c(1640, 1828))
  print(p)
}
```
